---
title: "C19 Analysis"
output: pdf_document
date: "2023-01-31"
---

## 1. Set up
Note: the following R packages are used for this analysis. Please refer to the session info at the end of the report.

```{r setup, include=TRUE, echo=TRUE}
library(readr)
library(tidyverse)
library(lubridate)
library(forecast)
library(patchwork)
```

## 2. Getting the Data
```{r getting_data, include=TRUE, echo=FALSE, warning=FALSE}
url_in <- 'https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/'
file_names <- c(
  "time_series_covid19_confirmed_global.csv",
  "time_series_covid19_deaths_global.csv",
  "time_series_covid19_confirmed_US.csv",  
  "time_series_covid19_deaths_US.csv"
  )
urls <- str_c(url_in, file_names)

#global_cases <- read_csv(urls[1])
#global_deaths <- read_csv(urls[2])
us_cases <- read_csv(urls[3])
us_deaths <- read_csv(urls[4])
```

## 3. Examing and Wrangling the Data

We'll convert the data from wide to long format and then join them together to have both cases and deaths data together.

```{r wrangling_data, include=TRUE}
us_cases_long <- us_cases %>% 
  select(-c(UID, iso2, iso3, code3, FIPS, Admin2, Lat, Long_, Combined_Key)) %>% 
  rename("province_state"="Province_State", "country_region"="Country_Region") %>% 
  pivot_longer(cols=-c(province_state, country_region), names_to='date', values_to='cases')

us_deaths_long <- us_deaths %>% 
  select(-c(UID, iso2, iso3, code3, FIPS, Admin2, Lat, Long_, Combined_Key)) %>% 
  rename("province_state"="Province_State", "country_region"="Country_Region", "population"="Population") %>% 
  pivot_longer(cols=-c(province_state, country_region, population), names_to='date', values_to='deaths') 
```

```{r joining_data, include=TRUE} 
us_cases_long$date <- as.Date(us_cases_long$date, format="%m/%d/%y")
us_deaths_long$date <- as.Date(us_deaths_long$date, format="%m/%d/%y")

florida_cases <- us_cases_long %>% 
  filter(province_state=="Florida" & date <= as_date('2020-12-31')) %>% 
  select(date, cases)
florida_deaths <- us_deaths_long %>% 
  filter(province_state=="Florida" & date <= as_date('2020-12-31')) %>% 
  select(date, deaths, population)

florida_cases <- florida_cases[1:345,]
florida_deaths <- florida_deaths[1:345,]

```

## 4. Analyze the Data
We'll use the data to answer the following questions:
- In general, does the number of cases follow the number of deaths trend? We'll use the state of Florida's data to answer this question. 
- How does the death/case rate differ across states?

We'll then build a model to predict number of deaths given number of cases.

First, let's plot the cases and deaths side by side:

```{r visualizing_data_1, include=TRUE} 
florida <- merge(florida_cases, florida_deaths, by="date")

case_plot <- florida %>% ggplot(aes(x=date, y=cases)) + geom_line()
deaths_plot <- florida %>% ggplot(aes(x=date, y=deaths)) + geom_line()

case_plot + deaths_plot

```
Clearly, they move together. Now let's put them in the same plot:

```{r visualizing_data_2, include=TRUE} 
ggplot(florida, aes(x=date)) +
         geom_line(aes(y=cases/30)) +
         geom_line(aes(y=deaths))

```
Again, there's a correlation between the two.

Next, let's build a model to predict deaths based on cases.
```{r modeling, include=TRUE}
lm(formula=deaths~cases, data=florida)
```


## 5. Conclusion
In this analysis, we looked at the relationship between cases and deaths and the visualization showed there's a correlation betweem them. We also built a linear regression model to predict deaths based on cases.

## 6. Session Info
```{r sessioninfo, include=TRUE, echo=TRUE}
sessionInfo()
```

